% ======================================
% MAIN MANUSCRIPT STRUCTURE (CLEANED)
% ======================================

\documentclass[11pt]{article}
\input{preamble}

\begin{document}

\title{Structured Orb Dynamics:\\ Unified Manuscript and Data Repository}
\author{Cassandra Perry}
\date{\today}

\maketitle

\begin{abstract}
This manuscript develops a unified geometric framework for interpreting unknown aerial motion using only observable trajectory structure. The approach combines stabilized image-plane reconstruction, curvature-based analysis, and a discrete motion-state model to describe how an object’s motion evolves over time, independent of sensor metadata or assumed physical mechanisms.

The framework links curvature, its temporal behavior, and associated kinematic features to four interpretable motion states—straight, turn, hover, and orb. These states are supported by a likelihood model that evaluates geometric consistency frame by frame while maintaining temporal coherence through a minimal set of transition assumptions. The result is a classifier that summarizes motion behavior in a way that is reproducible, dataset-agnostic, and grounded entirely in observable quantities.

Applied to publicly released infrared datasets such as PR-018, the method identifies sustained, smoothly varying curvature regimes that are not well described by classical motion categories alone. The model does not infer intent or mechanism; instead, it provides a transparent description of the motion’s geometric organization and the transitions between distinct dynamical regimes.

This framework offers a consistent basis for comparing trajectories across heterogeneous datasets and establishes a clear path for future work in geometric motion analysis where traditional metadata is incomplete or unavailable.
\end{abstract}

\tableofcontents
\clearpage

% ======================================
% PART I – INSTRUMENTATION & PIPELINE
% ======================================

\section*{Part I: Instrumentation and Reconstruction Pipeline}
\addcontentsline{toc}{section}{Part I: Instrumentation and Reconstruction Pipeline}
\input{manuscripts/Part_I_Instrumentation_Model/part1}

\clearpage

% ======================================
% PART II – THEORETICAL FRAMEWORK
% ======================================

\section*{Part II: Theoretical Framework}
\addcontentsline{toc}{section}{Part II: Theoretical Framework}
\label{sec:part2_intro}
\input{manuscripts/Part_II_Structured_Orb_System/part2}

\clearpage

% ======================================
% PART III – PR018 DEEP ANALYSIS
% ======================================

\section*{Part III: PR018 Deep Analysis}
\addcontentsline{toc}{section}{Part III: PR018 Deep Analysis}
\input{manuscripts/Part_III_PR018_Deep_Analysis/part3}

\clearpage

% ======================================
% PART IV – SUPPLEMENTARY GIMBAL ANALYSIS
% ======================================

\section*{Part IV: Supplementary GIMBAL Analysis}
\addcontentsline{toc}{section}{Part IV: Supplementary GIMBAL Analysis}
\input{manuscripts/Supplementary_GIMBAL_Analysis/gimbal}

\clearpage

\bibliographystyle{ieeetr}
\bibliography{references}

\end{document}